Taking Conditional Independence Seriously:\\A Latent Network Approach to Dyadic Analysis

### 2017-10-27 paper/4_app.tex
app edits 2

% In terms of out-of-sample performance, shown in Figure~\ref{fig:weeksauc}, the AME model performs markedly better than the GLM out-of-sample, lending credence to the possibility that these third-order dependencies cause spurious effects for regime type.  In terms of performance, the original Weeks' model does slightly better than a coin-toss overall, and fails to predict more than a couple of events (incorrectly) in almost one million observations.

@@

The replications we have undertaken are all from articles over the past fifteen years.  A more recent example is \citet{gibler:2017} which examines the onset of militarized disputes using capabilities, joint democracy, alliances, and power parity in a undirected dyadic study using logistic regression and dyad clustered standard errors.   In addition to this, Gibler shows that the long-standing relationship between the relative parity of capabilities and initiation of international conflict is almost completely mediated by the initial conditions for the members of the dyad when they joined the international system as sovereign members. This finding calls into question many IR theories about the role of balance in terms of generating international conflict \citep{organski:1958}.

% SM: we still adding a coefplot here?
% Insert coefplot here, without splines...\label{fig:giberrepl}
% %
% %\begin{table}
% %\begin{center}
% %\caption{Comparison of Gibler (2017) Model 6 results with AME results. This was run over a chain of $10,000 \text{BAZILLION}$ iterations.  \label{tab:gibme}}
% %\begin{tabular}{lrr} \toprule
% %& \multicolumn{2}{c}{$\hat{\beta}$ Estimates}\\ \cmidrule{2-3}
% %Variable & \texttt{logit} & \texttt{amen} \\ \midrule
% %Allied & 0.142&  0.023 \\
% %Joint Democracy &\bf -0.507&  0.045 \\
% %Peace Years \bf &-0.260& \bf -0.060 \\
% %Spline 1 &\bf -0.001& \bf 0.00 \\
% %Spline 2&\bf -0.000& \bf 0.00\\
% %Spline 3 &-0.000& \bf 0.00\\ 
% %Contiguity &\bf 2.412& \bf 0.640 \\
% %Parity &0.075&  -0.013 \\
% %Parity at entry year&\bf 0.868&  0.002 \\ 
% %Rivalry &\bf 2.031&\bf 0.721\\ 
% %Constant &\bf -5.526 & \bf  -2.581\\ \bottomrule
% %\end{tabular}\\
% %\end{center}
% %{\bf Note}: {\bf Bold} indicates conventional statistical significance at $p < 0.05$ or less in Gibler (2017). For comparative purposes, only, we have employed the same criterion to the results from the \texttt{amen} estimation. NEED TO BE FIXED if staying.
% %\end{table}

@@

First and foremost, many findings which emerge from models that do not take interdependencies into account lose their statistical significance when network effects are estimated via AME.\footnote{This should not be news---since the finding has long been in the theoretical literature, but given the state of current literature in international relations is still pertinent.} Not only are coefficients biased in the GLM approaches to the analysis of dyadic data, but they are often imprecisely measured, with poorly calibrated standard errors.  This means that significant testing (for better or worse) is compromised when network effects are ignored.

@@

Fifth, it is useful to suggest that a lot of what is \textit{known} about international relations based on studies of dyadic data should be taken with a grain of salt, awaiting a reassessment using methods that do not rely on an assumption that all the dyads are iid.

>>>>

@@

The replications we have undertaken are all from articles over the past fifteen years.  A more recent example is \citet{gibler:2017} which examines the onset of militarized disputes using capabilities, joint democracy, alliances, and power parity in a undirected dyadic study using logistic regression and dyad clustered standard errors. In addition, Gibler shows that the long-standing relationship between the relative parity of capabilities and initiation of international conflict is almost completely mediated by the initial conditions for the members of the dyad when they joined the international system as sovereign members. This finding calls into question many IR theories about the role of balance in terms of generating international conflict \citep{organski:1958}.

@@

First and foremost, many findings that emerge from models that do not take interdependencies into account lose their statistical significance when network effects are estimated via AME.\footnote{This should not be news---since the finding has long been in the theoretical literature, but given the state of current literature in international relations is still pertinent.} Not only are coefficients biased in the GLM approaches to the analysis of dyadic data, but they are often imprecisely measured, with poorly calibrated standard errors.  This means that significant testing (for better or worse) is compromised when network effects are ignored.

@@

% i thinks that's already pretty explicit
% Fifth, it is useful to suggest that a lot of what is \textit{known} about international relations based on studies of dyadic data should be taken with a grain of salt, awaiting a reassessment using methods that do not rely on an assumption that all the dyads are iid.
